---
output:
  bookdown::pdf_document2:
    toc : no
    keep_tex: no
    number_sections: yes
---

# Methods



## Projecting species distributions

We projected changes in biomass and potential catches of exploited species using a linked climate-fish-fisheries model called the dynamic bioclimate envelope model (DBEM). See Cheung *et al *(2009, 2016a, 2016b) for in-depth descriptions of the model. In brief, the DBEM has a horizontal spatial resolution of 0.5 o latitude x 0.5 o longitude for the sea surface and bottom, and simulates annual average abundance and catches of each modelled species. The DBEM uses spatially explicit outputs from Earth System Models, including temperature, oxygen level, salinity, surface advection, sea ice extent and net primary production. Sea bottom and surface temperature, oxygen and salinity are used for demersal and pelagic species, respectively. These outputs are then used to calculate an index of habitat suitability for each species in each spatial cell. Other information used to calculate habitat suitability includes bathymetry and specific habitats (coral reef, continental shelf, shelf slope, and seamounts). Changes in carrying capacity in each cell is assumed to be a function of the estimated habitat suitability, and net primary production in each cell [@Cheung.2016; @Cheung.2009; @Cheung.2016fx]. 

The model simulates the net changes in abundance in each spatial cell based on logistic population growth, fishing mortality, and movement and dispersal of adults and larvae modelled through advection–diffusion–reaction equations. Biomass is calculated from abundance by using a characteristic weight representing the average mass of an individual in the cell. The model simulates how changes in temperature and oxygen content would affect the growth of the individual using a submodel derived from a generalized von Bertalanffy growth function. 

### Rellocating fishing effort

Fishing intensity was assumed in the fisheries scenarios, represented by fishing mortality rates (*F*) relative to the *F* required to achieve maximum sustainable yield (i.e., $F_{MSY}$). As the model assumes logistic population growth, following the derivation from a simple surplus production model, FMSY is approximately equal to half of the intrinsic growth rate of each species. 

To reallocate fishing in the DBEM from protected grids to surrounding [@PalaciosAbrantes.2023c8m], we took the *area protected* and divided by the *total area* of each grid cell that contain any protection. We then reallocate that protected area proportion to the surrounding cells in form of fishing effort (*F*):

$$prop = 1 + \frac{area_{protected}/area_{total}}{n_{surrounding}}$$

where $n_{surrounding}$ is the number of grid cells surrounding an MPA.

Thus, for example, if status = protected, prop = 0 while if status = unprotected then prop = 1. Moreover, if there are 4 cells surrounding the protected one cell, then prop = 1.25, if there are 2 surrounding cells then prop = 1.5,if there are 10 surrounding cells then prop = 1.1. If, for example, the protected cell is only partially covered by a protected area, say 50%, and there are 4 surrounding cells, then prop = .5 for the protected cell, and prop = 1.125 (1 + .5/4) for the surrounding cells.

Based on the computed prop, we re-estimated fishing mortality (*F*) (i.e., reallocate fishing effort) in the *surrounding* cells as:

$$\hat{F}= F*prop$$

where $\hat{F}$ is the fishing mortality adjusted for the ‘spill-over’ of fishing effort from the *protected* grid cell. If the cell is not protected nor surrounding a protected grid cell then $\hat{F}$ = *F*, that is, fishing mortality will not be adjusted for the no-take MPA effects. 


### Climate change impacts on MPAs

To explore the impacts of climate change in the MPAs, we evaluated the percentage change in biomass and MCP in each grid cell ($ΔB_i$) between two time periods representing the present time ($B_p$ - 1995–2014) and the mid of the 21st century ($B_m$ 2030–2049) to reduce climate variability.

$ΔB_i = \frac{B_p − B_m}{B_m}*100$ 

where $B_m$ is the biomass at the mid of the 21st century, $B_p$ is the biomass at present time, and *i* represents every grid cell in the study area. Note that for the aggregated results we sum the biomass of all species in each grid cell prior to estimate $ΔB_i$

### Climate change models and scenarios

The DBEM was forced with projections from three new-generation Earth System Models (ESMs) from Phase 6 of the Coupled Model Intercomparison Project (CMIP6). The ESMs included were the Geophysical Fluid Dynamics Laboratory (GFDL)-ESM4 [@Dunne.2013], the Institut Pierre-Simon Laplace (IPSL)-CM6A-LR [@Boucher.2020], and the Max Planck Institute Earth System Model (MPI)-ESM1.2 [@Gutjahr.2019]. Climate change projections followed two contrasting scenarios according to the shared socio-economic pathways (SSPs) and Representative Concentration Pathways (RCPs): SSP1-RCP2.6 (SSP1-2.6) and SSP5-RCP8.5 (SSP5-2.6) scenarios [@Gutschow.2021]. The SSP1-2.6 and SSP5-8.5 scenarios assume “strong mitigation” and “no mitigation” with radiative forcing stabilized at 2.6 W/m² and 8.5 W/m², respectively, by the end of the 21st century.

## Study area

We gridded the area around the MPAs to a spatial resolution of $0.5^o$ latitude x $0.5^o$ longitude to match the DBEM resolution. We classified the grid cells into protected, surrounding, and open waters (i.e., not protected). Protected areas include cells that contain an MPA polygon (cells that fully or partially contain an MPA) while cells that were immediately adjacent to an MPA cell were classified as surrounding waters. All other ocean cells were considered unprotected (Figure \@ref(fig:dbemgridFig)). We limit the study area (i.e., analysis area) to 200 km (~ $2^0$) around the border of the MPAs.



```{r dbemgridFig, eval = T, echo = T, fig.width = 4, fig.height = 4,fig.keep='high', echo=FALSE, fig.cap="Coco MPA in blue and AMMB area in red over DBEM grid cells (light blue) included in the analysis"}

# Call the image
img <- readPNG("./figures/study_area.png")

# Print the image
grid.raster(img)
```


## Conservation scenarios

The current project considers three different scenarios which differ in the amount and location available for fishing. The scenarios were build in collaboration with MarViva and range from a conservation inclined scenario (S1) to a more fisheries inclined scenario (S2). Below we detail each scenario:

### Scenario S1 (Conservation inclined)

-	PNC, 100% protected
-	AMM, *F* = $\frac{1}{2}MSY$
-	Everything else F = $F_{MSY}$
-	Reallocate fishing effort from PNC and AMM to the grid cells surrounding AMM


```{r ScenOneFig, eval = T, echo = T, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Scenario 1. Prop = proportion of area that is open to fishing."}

# Call the image
img <- readPNG("./figures/scenario_one.png")

# Print the image
grid.raster(img)
```


### Scenario S2 (Fishing inclined):
-	PNC, 100% protected
-	AMM, *F* = $F_{MSY}$
-	Everything else *F* = $F_{MSY}*1.5$
-	Fishing from PNC gets reallocated to cells surrounding PNC (which are AMM cells)

```{r ScenTwoFig, eval = T, echo = T, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Scenario 2. Prop = proportion of area that is open to fishing"}

# Call the image
img <- readPNG("./figures/scenario_two.png")

# Print the image
grid.raster(img)
```

 
### Scenario S3 (IUU inclined)
-	PNC, 90% protected (Allows for 10% fishing as IUU)
-	AMM, *F* = $F_{MSY}*1.5$
-	Everything else *F* = $F_{MSY}*1.5$
-	fishing from PNC gets reallocated to cells surrounding PNC (which are AMM cells)


```{r ScenThreeFig, eval = T, echo = T, fig.width = 5, fig.height = 5, echo=FALSE, fig.cap="Scenario 3. Prop = proportion of area that is open to fishing"}

# Call the image
img <- readPNG("./figures/scenario_one.png")

# Print the image
grid.raster(img)
```
 
## Species

The current analysis explored the impacts of climate change on 38 species that occur in the Eastern Tropical Pacific ([See Species List on GitHub repository](https://github.com/jepa/cocos_mpa/blob/main/data/species/project_species_list.csv)). We grouped the species into four different commercial groups according to [The Sea Around Us](http://www.seaaroundus.org/): perch-likes (*n* = 5), sharks and rays (*n* = 8), tunas and billfishes (*n* = 9) and other fishes (*n* = 2).

```{r spp_list, eval = F, echo = F}
read_csv("../data/species/project_species_list.csv") %>% 
  kbl(caption = "Table 1. List of species included in this report")
```


## Code and Data

All of the data analysis and statistics were completed with the statistical software R version 4.3.1 (2023-06-16) and can be found at github.com/jepa/cocos_mpa. Data for this project can be downloaded from [this repository](https://ubcca-my.sharepoint.com/:f:/g/personal/jepa88_student_ubc_ca/ErdLryE-_L1IjeSlVoPBSxYBIPFX-DU5ztjnO8DWBfRRqg?email=t.clarke%40oceans.ubc.ca&e=u73Rkb)[^1].


[^1]: Currently only Tayler Clarke with her *UBC* email has access to the data. We are happy to give access to other people, we only require the name and email. 


## Caveats

There are a few caveats to this report that are important to take into consideration when reading the results. First, the DBEM is a mechanistic species distribution model, thus, it does not take into consideration species interactions, species potential evolutionary adaptation through space and time, nor any other human-induced effect (e.g., mining, shipping, pollution) [@Gaston.1996; @Peterson.2012]. Second, as any mathematical model, the DBEM is subjected to the original input data in terms of quality, quantity, and scale (See report Data folder, *distributions*). The outputs from the ESMs are produced at a 1 degree lat x long resolution and further downscaled to 0.5 degrees to match the DBEM resolution. Yet, such a resolution might not be coarse enough to capture local (coastal) processes that are affecting the ecosystem around PNC and AMM. Such elements in combination with the small size of the MPA, especially cocos which represents 15 complete grid-cells and 9 partial cells, creates a substantial level of uncertainty in the projections. Third, fishing effort is redistributed proportionally around the grid-cells that are closed to fishing [@PalaciosAbrantes.2023c8m]. In reality, it is possible that such effort is not proportionally distributed around the no-take MPA but rather localized in some areas, or not re-distributed at all. Other modelling exercises that incorporate different mechanisms in which fishing effort is re-allocated around a no-take MPA could help illuminate how social and economic factors related to fishing may affect the contribution of MPAs under climate change. For these reasons, the results presented here, even at their most aggregated level, should be interpreted as an example of what could happen to the region, rather than definite predictions of what will happen to each stock of the Eastern Tropical Pacific under climate change. Replicating the analysis here with finer-scale models (e.g., Regional Ocean Modeling System - ROMS) and data (if existent) would help validate the DBEM while also producing more reliable results. 

