---
title: "pred_dbem"
author: "Sarah Roberts & Juliano Palacios Abrantes"
date: "2023-10-24"
output: html_document
---

```{r setup, include = F}
source(here::here('./functions/function_list.R')) #I took these from mpa_c_toe

load_pkg(c(
  # For grid estimation and species selection
  "spatialEco","tidyverse", "geosphere","raster","units","matrixStats","sf","rmapshaper", "igraph",
  # For collaborative ease
  "here",
  "readr",
  "janitor"
  ))

# For grid e
# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

select <- dplyr::select

```

# Load data

```{r}

# Scenario sf
ammb_sf <- st_read(MyFunctions::my_path("D","spatial/ammb","AMM_Bicentenario_CostaRica.shp")) %>% 
  st_transform(4326)

st_crs(ammb_sf) <-4326

coco_sf <- st_read(MyFunctions::my_path("D","spatial/coco","PN_IslaCoco_CostaRica.shp")) %>% 
  st_transform(4326)
st_crs(coco_sf) <-4326



# Scenario grids
scen_grid <- 
  bind_rows(
  read.csv("../data/scenario1.csv") %>% mutate(scen = "S1"),
  read.csv("../data/scenario2.csv") %>%  mutate(scen = "S2"), # Why scenario 2 has prop2 and prop3? using prop for now
  read.csv("../data/scenario3.csv") %>% mutate(scen = "S3")
  ) %>% 
  select(index = Seq, scen, Lat,Lon,prop) %>% 
  clean_names() %>% 
  mutate(
    status = ifelse(prop == 1,"open",
                    ifelse(prop <= 1,"protected",
                           ifelse(prop >= 1,"surrounding","outside")
                    )
    )
  )


# Load DBEM runs

# dbem_runs <- load_dbem("C6GFDL85F1MPAS1", cat = "Abd")


# Catch
dbem_runs_c <- bind_rows(
  lapply(c("C6GFDL85F1MPAS1",
           "C6GFDL85F1MPAS2",
           "C6GFDL85F1MPAS3"), load_dbem, cat = "Catch")
)%>% 
  mutate(variable = "Catch") %>% 
  left_join(scen_grid,
            by = c("index","scen")
            )

unique(dbem_runs_c$scen)
unique(dbem_runs_c$status)

# Abundance
dbem_runs_a <- bind_rows(
  lapply(c("C6GFDL85F1MPAS1",
           "C6GFDL85F1MPAS2",
           "C6GFDL85F1MPAS3"), load_dbem, cat = "Abd")
) %>% 
  mutate(variable = "Abd") %>% 
  left_join(scen_grid,
            by = c("index","scen")
            )

unique(dbem_runs_a$scen)
unique(dbem_runs_a$status)


```


# Preeliminary results

## numbers exploration

There is definitely something wrong with the runs 

- Scen 1 has no abundance on surrounding grids and no catch on surrounding 
- Scen 2 has no abundance on oppen grids and no catch data on oppen nor protected
- Scen 3 is the only one working

```{r}

bind_rows(dbem_runs_a,dbem_runs_c) %>% 
  group_by(scen,variable,status) %>% 
  summarise(total_value = sum(value, na.rm = T),
            mean_value = mean(value, na.rm = T)
            ) %>% 
  View()
```


## Map results

```{r}

dbem_runs %>% 
  group_by(index,year,scen) %>% 
  summarise(
    total_value = sum(value,na.rm = T)
  ) %>% 
  filter(year %in% c(1951, 1980, 2010, 2040, 2070,2099),
         lat > 1,
         lat < 8,
         lon < - 83) %>% 
  # View()
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = total_value,
      color = total_value
    )
  ) +
  scale_fill_viridis_b() +
  scale_color_viridis_b() +
  facet_grid(year~scen) +
  geom_sf(data = ammb_sf, aes(), fill = "transparent", color = "red", size = 2) +
  geom_sf(data = coco_sf, aes(), fill = "transparent", color = "black", size = 2) +
  theme(legend.position = "bottom")

ggsave(filename = "../results/catch_scen.png",plot = last_plot(), height = 14)

```
## Linear change by grid

```{r}

dbem_runs %>% 
  group_by(index,year,scen) %>% 
  summarise(
    total_value = sum(value,na.rm = T)
  ) %>% 
  group_by(year,status,scen) %>% 
  summarise(
    total_value = sum(total_value,na.rm = T)
  ) %>% 
  # filter(scen == "S1") %>% 
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = total_value,
      color = scen
    )
  ) +
  facet_wrap(~status)
  
ggsave(filename = "../results/catch_scen_lin.png",plot = last_plot())

```
## Liner change by grid

This is weighted method of abundance by grid type/number of grids

```{r}

# Estimate ammount of grids
n_status <- scen_grid %>% 
  group_by(status,scen) %>% 
  tally()

# Show result
dbem_runs %>% 
  group_by(index,year,scen) %>% 
  summarise(
    total_value = sum(value,na.rm = T)
  ) %>% 
  group_by(year,status,scen) %>% 
  summarise(
    total_value = sum(total_value,na.rm = T)
  ) %>% 
  left_join(n_status,
            by = c("status","scen")
            ) %>% 
  mutate(value_wgt = total_value/n) %>% 
  group_by(status,scen) %>% 
  mutate(
    value_wgt_rm = round(zoo::rollmean(
      value_wgt,
      k = 10, 
      fill = NA, 
      align = "left")
    )
  ) %>% 
  filter(!is.na(value_wgt_rm)) %>% 
  View()
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = value_wgt, # does not like running mean on catch
      color = status,
      linetype = scen
    )
  )
  

```