---
title: "pred_dbem"
author: "Sarah Roberts & Juliano Palacios Abrantes"
date: "2023-08-24"
output: html_document
---

```{r setup, include = F}
source(here::here('./functions/function_list.R')) #I took these from mpa_c_toe

load_pkg(c(
  # For grid estimation and species selection
  "spatialEco","tidyverse", "geosphere","raster","units","matrixStats","sf","rmapshaper", "igraph",
  # For collaborative ease
  "here",
  "readr"
  ))

# For grid e
# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

select <- dplyr::select

```


# 1 Load data and Intersect
##1.1 Load in the mpa data and check crs information
```{r}
ammb <- read_sf((here("data/spatial/ammb/AMM_Bicentenario_CostaRica.shp")))
plot(ammb)

#get out coordinate reference system information 
st_crs(ammb)
st_crs(ammb)$epsg

crs <- st_crs(ammb) #this is a costa rica specific epsg with WGS 84 datum


coco <- read_sf((here("data/spatial/coco/PN_IslaCoco_CostaRica.shp")))
plot(coco)
st_crs(coco)
st_crs(coco)$epsg #this is in a UTM coordinate reference system for the costa rica area 

#lets have them both in the WGS one 

coco <- st_transform(coco, crs)
st_crs(coco)

#plot them both to make sure it looks correct 
ggplot() + geom_sf(data = ammb, colour = "blue") + geom_sf(data = coco, colour = "red") 
```

## 1.2 Load in grid data
```{r}
grid <- read_sf(here("./data/spatial/grid/worldsq_EA.shp"))
grid <- grid %>% filter(Lat < 10 & Lat > 0 & Lon > -92 & Lon < -82)

grid <- st_transform(grid, crs = st_crs(crs)) %>% 
  filter(PWater > 0)

ggplot() + geom_sf(data = ammb, colour = "blue")+ geom_sf(data = coco, colour = "red") 

ggplot() + geom_sf(data = grid) + geom_sf(data = ammb, aes(colour = "ammb"), fill = NA) + geom_sf(data = coco, aes(colour = "coco"), fill = NA) +
scale_color_manual(values = c("darkred", "blue"))
```

## 1.3 Identify grids that are MPAs
```{r}

#coco 
grid_intersection <- st_intersection(grid, coco) %>%
  group_by(Seq) %>%
  summarise(geom = st_union(geometry)) %>%
  mutate(geom = st_sfc(geom),
         area = st_area(geom))
grid_intersection$status <- "protected_coco" #lets us know that those grid cells have some level of MPA within them 

#join this back to the big grid (right now we just have the gridIDs that are overlapping an MPA)
grid_intersection <- as.data.frame(grid_intersection)

grid2 <- as.data.frame(grid)
grid2 <- left_join(grid2, grid_intersection, by = "Seq")
grid2$protected_area_coco_m2  <- as.numeric(grid2$area)
grid2 <- grid2 %>% dplyr::select(-geom)
grid2 <- st_as_sf(grid2)

ggplot() + geom_sf(data = grid2, aes(fill = protected_area_coco_m2)) + geom_sf(data = ammb, aes(colour = "ammb"), fill = NA) + geom_sf(data = coco, aes(colour = "coco"), fill = NA) +
scale_color_manual(values = c("darkred", "blue"))

#ammb 
grid_intersection2 <- st_intersection(grid, ammb) %>%
  group_by(Seq) %>%
  summarise(geom = st_union(geometry)) %>%
  mutate(geom = st_sfc(geom),
         area = st_area(geom))

grid_intersection2$status <- "protected_ammb" #lets us know that those grid cells have some level of MPA within them 

#join this back to the big grid (right now we just have the gridIDs that are overlapping an MPA)
grid_intersection2 <- as.data.frame(grid_intersection2)

grid3 <- as.data.frame(grid2)
grid3 <- left_join(grid3, grid_intersection2, by = "Seq")
grid3$protected_area_ammb_m2  <- as.numeric(grid3$area.y)
grid3 <- grid3 %>% dplyr::select(-geom)
grid3 <- st_as_sf(grid3)

ggplot() + geom_sf(data = grid3, aes(fill = protected_area_ammb_m2)) + geom_sf(data = ammb, aes(colour = "ammb"), fill = NA) + geom_sf(data = coco, aes(colour = "coco"), fill = NA) +
scale_color_manual(values = c("darkred", "blue"))

grid3 <- grid3 %>% rename(status_coco = "status.x", status_ammb = "status.y") %>% select(-area.x, -area.y)

```
##1.4 Estimate area

```{r}

grid3$grid_area <- st_area(grid3)
grid3 <- grid3 %>% drop_units()

grid3$prop_coco <- 1
grid3$prop_ammb <- 1
grid3 <-
  grid3 %>% 
  mutate(
    prop_coco = ifelse(status_coco == "protected_coco", (protected_area_coco_m2/grid_area), prop_coco),
    prop_ammb = ifelse(status_ammb == "protected_ammb", (protected_area_ammb_m2/grid_area), prop_ammb)
  ) #this is the proportion of the grid cell covered by each mpa 


```

## 1.5 Find surrounding grid cells 
Because ammb is surrounding cocos (i.e cocos is completely within ammb), we will find both the surrounding grid cells for ammb and cocos. The cocos effort and ammb effort will be reallocated to those cells depending on the scenario.

```{r}
#start with coco
grid_coco <- grid3 %>% 
  rename(status = "status_coco", protected_area_m2 = "protected_area_coco_m2") %>% 
  select(OBJECTID, Seq, Lat, Lon, geometry, status, protected_area_m2, grid_area) %>% 
  mutate(mpa = "cocos") #use to calculate size of the Mpa 

grid_coco_mpa <- grid_coco %>% filter(status == "protected_coco")
grid_coco_mpa$status <- "protected"


grid_coco_no <- grid_coco %>% filter(is.na(status))

ggplot() + geom_sf(grid_coco_mpa, mapping = aes(colour = as.factor(mpa)))

#create buffer
buffer <- st_buffer(grid_coco_mpa, 54500)

intr <- st_intersection(grid_coco_no, buffer)
intr_seqs <- c(intr$Seq)
grid_coco_surround <- grid_coco_no %>% filter(Seq %in% intr_seqs)



grid_coco_surround <- grid_coco_surround %>% mutate(surrounding_coco ="surrounding_coco")


ggplot() + geom_sf(data = grid_coco_surround, aes(fill = surrounding_coco)) + geom_sf(data = grid_coco_mpa, aes(fill = status))

#now ammb 

grid_ammb <- grid3 %>% 
  rename(status = "status_ammb", protected_area_m2 = "protected_area_ammb_m2") %>% 
  select(OBJECTID, Seq, Lat, Lon, geometry, status, protected_area_m2, grid_area) %>% 
  mutate(mpa = "ammb") #use to calculate size of the Mpa 

grid_ammb_mpa <- grid_ammb %>% filter(status == "protected_ammb")
grid_ammb_mpa$status <- "protected"


grid_ammb_no <- grid_ammb %>% filter(is.na(status))

ggplot() + geom_sf(grid_ammb_mpa, mapping = aes(colour = as.factor(mpa)))

#create buffer
buffer <- st_buffer(grid_ammb_mpa, 54500)

intr <- st_intersection(grid_ammb_no, buffer)
intr_seqs <- c(intr$Seq)
grid_ammb_surround <- grid_ammb_no %>% filter(Seq %in% intr_seqs)



grid_ammb_surround <- grid_ammb_surround %>% mutate(surrounding_ammb ="surrounding_ammb")


ggplot() + geom_sf(data = grid_ammb_surround, aes(fill = surrounding_ammb)) + geom_sf(data = grid_ammb_mpa, aes(fill = status))


#create one big dataframe - there are a few that are both surrounding ammb and coco so keep that column separate for now

grid_ammb_surround <- grid_ammb_surround %>% select(Seq, surrounding_ammb) %>% st_drop_geometry()
grid_coco_surround <- grid_coco_surround %>% select(Seq, surrounding_coco) %>% st_drop_geometry()
grid_use <- left_join(grid3, grid_ammb_surround, by = "Seq")
grid_use <- left_join(grid_use, grid_coco_surround, by = "Seq")

grid_use <- grid_use %>% mutate(surrounding_ammb = replace(surrounding_ammb, which(status_coco == "protected_coco"), NA)) #we don't need these to be considered surrounding 


#total proportion 
#this will help allocate fishing effort to the surrounding areas. 
tot_cocos_area = sum(grid_coco_mpa$grid_area)
tot_cocos_mpa = sum(grid_coco_mpa$protected_area_m2)
prop_cocos = tot_cocos_mpa/tot_cocos_area

tot_ammb_area = sum(grid_ammb_mpa$grid_area)
tot_ammb_mpa = sum(grid_ammb_mpa$protected_area_m2)
prop_ammb = tot_ammb_mpa/tot_ammb_area

tot_ammb_or_cocos_area <- grid_use %>% filter(status_coco %in% "protected_coco" | status_ammb %in% "protected_ammb")
tot_ammb_or_cocos_area = sum(tot_ammb_or_cocos_area$grid_area) #this is necesairy so we don't double count protected cells in scenario 1




```

#Reallocate 
There are three ways to reallocate

1. Based on area: To reallocate fishing from protected, we take the area protected/total area of grid cells that contain a protected area and reallocate that protected area proportion to the surrounding cells. This is the most accurate and least conservative (i.e least amount of fishing)
  (area covered by mpa*protection level/total area of grid cells that contain a protected area)/n surrounding grid cells + 1 (or 1.5)

2. Based on grid cell #: (Number of grid cells containing any level of protection*protection level/n surrounding grid cells) + 1.  This is more conservative and assumes if any amount of a grid cells is protected then any fishing that would have happened in that grid cell gets reallocated. It is less accurate to calculate area of grid cells for the entire globe so we did it this way. This is the most conservative (i.e allows the most fishing)

3. Based on edges: Reallocate only edge grid cells. This was done for North Sea because the MPAs were small and random so most of them only contained edge grid cells. Also, there were no grid cells that were partially covered. This one just doesn't make sense for the way these MPAs area designed, but it is in between 1 and 2 for prop allowed to fish 

#2 Reallocate based on area
## 2.1 Scenario 1 
(Conservation included)
PNC, 100% MPA
AMM, FMSY = 1/2MSY
Everything else F = MSY
  Reallocate the fishing from PNC and AMM to the grid cells surrounding AMM

Prop is going to be proportion of the grid cell that fishes

1. calculate total area covered by PNC, AMM 

2. For grid cells with both PNC and AMM, prop in coco*fishing allowed + prop in ammb*fishing allowed =         .5*0 + .5*.5 = .25 for scenario 1
3. Set PNC prop to 1-prop in coco - for grid cells that only intersect PNC

4. Set AMM prop to .5*ammb prop of grid cell covered - for only intersect AMMB

5. Redistribute total PNC area and 1/2 AMM area to surrounding cells 
  calculate proportion to reallocate: 
    prop_allocation = (area covered by PNC + (area covered by AMM)*AMM fishing allowed)/area of grid cells containing AMM or PNC <- needs to be or, or else will double count grid cells
    
  Redistribute that to surrounding grid cells area of surrounding grid cells + 1 
    = (prop_allocation/# surrounding grid cells) + 1

```{r}
#grid cells that overlap both cocos and ammb 
grid_sc1 <- grid_use %>%  mutate(prop = if_else(status_coco %in% "protected_coco" & status_ammb %in% "protected_ammb", prop_coco*0 + prop_ammb*.5,1)) #%in% important here because of NAs

#grid cells that are just ammb 
grid_sc1 <- grid_sc1 %>%  mutate(prop = if_else(status_ammb %in% "protected_ammb" & !status_coco %in% "protected_coco", 1-(prop_ammb*.5), prop))

#grid cells that are just coco 
grid_sc1 <- grid_sc1 %>%  mutate(prop = if_else(!status_ammb %in% "protected_ammb" & status_coco %in% "protected_coco", 1-prop_coco, prop)) #these should all be 0

#grid cells surrounding ammb (and coco)
n_surround = grid_sc1 %>% st_drop_geometry %>% filter(surrounding_ammb == "surrounding_ammb") %>% tally()

#protected proportion to reallocate (calculated as area of each mpa/area of grid cells containing mpa)
prop_allocate = (tot_ammb_mpa*.5 + tot_cocos_mpa)/tot_ammb_or_cocos_area

#allocate surrounding cells
grid_sc1 <- grid_sc1 %>%  mutate(prop = if_else(surrounding_ammb %in% "surrounding_ammb",(prop_allocate/n_surround$n) +1 , prop))

ggplot() + geom_sf(data = subset(grid_sc1, grid_sc1$prop != 1), aes(fill = prop)) + geom_sf(data = ammb, aes(colour = "ammb"), fill = NA) + geom_sf(data = coco, aes(colour = "coco"), fill = NA) +
scale_color_manual(values = c("darkred", "blue"))

```

##2.2 Scenario 2
(Fishing inclined)
- PNC, 100% MPA original
- AMM, FMSY
- Everything else F = 1.5MSY

fishing from PNC gets reallocated to cells surrounding PNC (which are AMM cells)
Prop is going to be proportion of the grid cell that fishes

1. calculate total area covered by PNC

2. Set PNC prop to 1-prop in coco - for grid cells that only intersect PNC, set AMM to 1

3. Redistribute PNC area protected to AMM cells that surround PNC 
   area covered by PNC*PNC fishing allowed/area grid cells containing PNC 
   + 1 

4. Set everything else to 1.5 

```{r}

#grid cells that overlap coco 
grid_sc2 <- grid_use %>%  mutate(prop = if_else(status_coco %in% "protected_coco", 1-prop_coco, 1)) #this means that fishing decreases by proportion of the cell covered by coco

#Redistribute PNC area fished to AMM cells that surround PNC 
#grid cells surrounding coco
n_surround = grid_sc2 %>% st_drop_geometry %>% filter(surrounding_coco == "surrounding_coco") %>% tally()

#protected proportion to reallocate (calculated as area covered by coco/area of grid cells containing coco mpa)
prop_allocate = tot_cocos_mpa/tot_cocos_area

#allocate surrounding cells
grid_sc2 <- grid_sc2 %>%  mutate(prop = if_else(surrounding_coco %in% "surrounding_coco",(prop_allocate/n_surround$n) +1 , prop))

#Set everything else to 1.5 
grid_sc2 <- grid_sc2 %>%  mutate(prop = if_else(status_ammb %in% "protected_ammb" | status_coco %in% "protected_coco" | surrounding_coco %in% "surrounding_coco", prop, 1.5))

ggplot() + geom_sf(data = grid_sc2, aes(fill = prop)) + geom_sf(data = ammb, aes(colour = "ammb"), fill = NA) + geom_sf(data = coco, aes(colour = "coco"), fill = NA) +
scale_color_manual(values = c("darkred", "blue"))
```

##2.3 Scenario 3 
(IUU inclined)
- PNC, 80%  (20% allows fishing)
- AMM, FMSY1.5
- Everything else F = 1.5MSY

fishing from PNC gets reallocated to cells surrounding PNC (which are AMM cells)
Prop is going to be proportion of the grid cell that fishes


1. Set PNC prop to 1-(prop in coco*.80) - for grid cells that only intersect PNC

2. Redistribute PNC area*prop protected to AMM cells that surround PNC 

3. Set everything else to 1.5 

```{r}

#grid cells that overlap coco 
grid_sc3 <- grid_use %>%  mutate(prop = if_else(status_coco %in% "protected_coco", 1-prop_coco*.80, 1.5)) #this means that fishing decreases by proportion of the cell covered by coco - so if it is 100% covered by coco then prop should be .1 open to fishing

#Redistribute PNC area fished to AMM cells that surround PNC 
#grid cells surrounding coco
n_surround = grid_sc3 %>% st_drop_geometry %>% filter(surrounding_coco == "surrounding_coco") %>% tally()

#protected proportion to reallocate (calculated as area covered by coco*.8/area of grid cells containing coco mpa)
prop_allocate = (tot_cocos_mpa*.8)/tot_cocos_area

#allocate surrounding cells
grid_sc3 <- grid_sc3 %>%  mutate(prop = if_else(surrounding_coco %in% "surrounding_coco",(prop_allocate/n_surround$n) +1.5 , prop)) #changed to 1.5 

ggplot() + geom_sf(data = subset(grid_sc3, grid_sc3$prop !=1.5), aes(fill = prop)) + geom_sf(data = ammb, aes(colour = "ammb"), fill = NA) + geom_sf(data = coco, aes(colour = "coco"), fill = NA) +
scale_color_manual(values = c("darkred", "blue"))

```


#3 Reallocate based on number of grid cells 
To test the sensitivity of each reallocation method, take Scenario 2 and reallocate based on number of grid cells (instead of proportion area protected). Scenario 2 is the most simple. 

##3.2 Scenario 2 
(Fishing inclined)
- PNC, 100% MPA original
- AMM, FMSY
- Everything else F = 1.5MSY

```{r}

#grid cells that overlap coco 
grid_sc2_2 <- grid_use %>%  mutate(prop = if_else(status_coco %in% "protected_coco", 1-prop_coco, 1)) #this means that fishing decreases by proportion of the cell covered by coco

#Redistribute PNC area fished to AMM cells that surround PNC 
#grid cells surrounding coco
n_surround = grid_sc2_2 %>% st_drop_geometry %>% filter(surrounding_coco == "surrounding_coco") %>% tally()

n_coco = grid_sc2_2 %>% st_drop_geometry %>% filter(status_coco == "protected_coco") %>% tally()

#allocate surrounding cells
grid_sc2_2 <- grid_sc2_2 %>%  mutate(prop = if_else(surrounding_coco %in% "surrounding_coco",(n_coco$n/n_surround$n) +1 , prop))

#Set everything else to 1.5 
grid_sc2_2 <- grid_sc2_2 %>%  mutate(prop = if_else(status_ammb %in% "protected_ammb" | status_coco %in% "protected_coco" | surrounding_coco %in% "surrounding_coco", prop, 1.5))

#there are 24 coco cells and 24 surrounding, so they all get allocated 

grid_sc2$prop2 <- grid_sc2_2$prop

ggplot() + geom_sf(data = grid_sc2, aes(fill = prop2)) + geom_sf(data = ammb, aes(colour = "ammb"), fill = NA) + geom_sf(data = coco, aes(colour = "coco"), fill = NA) +
scale_color_manual(values = c("darkred", "blue"))
```

#4 Reallocate based on outer cells only
To test the sensitivity of each reallocation method, take Scenario 2 and reallocate only outer grid cells. Scenario 2 is the most simple. I think the most meaningful way of doing this would be to reallocate based on grid cells (instead of area) and only do the outer grid cells (16 vs. 24) 

##3.2 Scenario 2 
(Fishing inclined)
- PNC, 100% MPA original
- AMM, FMSY
- Everything else F = 1.5MSY

```{r}

#grid cells that overlap coco 
grid_sc2_3 <- grid_use %>%  mutate(prop = if_else(status_coco %in% "protected_coco", 1-prop_coco, 1)) #this means that fishing decreases by proportion of the cell covered by coco

#Redistribute PNC area fished to AMM cells that surround PNC 
#grid cells surrounding coco
n_surround = grid_sc2_3 %>% st_drop_geometry %>% filter(surrounding_coco == "surrounding_coco") %>% tally()

n_coco = 16

#allocate surrounding cells
grid_sc2_3 <- grid_sc2_3 %>%  mutate(prop = if_else(surrounding_coco %in% "surrounding_coco",(n_coco/n_surround$n) +1 , prop))

#Set everything else to 1.5 
grid_sc2_3 <- grid_sc2_3 %>%  mutate(prop = if_else(status_ammb %in% "protected_ammb" | status_coco %in% "protected_coco" | surrounding_coco %in% "surrounding_coco", prop, 1.5))

#there are 24 coco cells and 24 surrounding, so they all get allocated 

grid_sc2$prop3 <- grid_sc2_3$prop

ggplot() + geom_sf(data = grid_sc2, aes(fill = prop3)) + geom_sf(data = ammb, aes(colour = "ammb"), fill = NA) + geom_sf(data = coco, aes(colour = "coco"), fill = NA) +
scale_color_manual(values = c("darkred", "blue"))
```

### 2.4 Clean up the dataset 

```{r}
names <- c("Seq", "Lat", "Lon", "prop")

sc1 <- grid_sc1 %>% st_drop_geometry() %>% dplyr::select(names)
sc3 <- grid_sc3 %>% st_drop_geometry() %>% dplyr::select(names)
sc2 <- grid_sc2 %>% st_drop_geometry() %>% dplyr::select(Seq, Lat, Lon, prop, prop2, prop3)


write.csv(sc1, "data/scenario1.csv")
write.csv(sc2, "data/scenario2.csv")
write.csv(sc3, "data/scenario3.csv")

```

# Prepare data for DBEM

```{r}

grid_to_dbem <- function(scenario){
  
  scenario_grid <- read_csv(paste0(here::here("data/"),scenario,".csv"))
  
  mpa_index <- scenario_grid %>% 
    rename(index = Seq) %>% 
    right_join(dbem_coords,
               by ="index") %>% 
  #   # Double check matched DBEM grid (it does)
  #   mutate(
  #     status = ifelse(prop == 1,"open",
  #                     ifelse(prop <= 1,"protected",
  #                            ifelse(prop >= 1,"surrouding","outside")
  #                     )
  #     )
  #   )
  # 
  # ggplot(mpa_index) +
  #   geom_tile(
  #     aes(
  #       x = lon,
  #       y = lat,
  #       fill = prop
  #     )
  #   ) +
  #   MyFunctions::my_land_map()
    
  ## Fix NAs for DBEM 
    arrange(index) %>% 
    mutate(prop = ifelse(is.na(prop),1,prop)) %>% 
    select(prop)
  
name <- paste0(here("data",scenario),".txt")
write.table(mpa_index, file=name, sep="\t", col.names = F, row.names = F)

}



# Run protocol
# Get DBEM coords
dbem_coords <- MyFunctions::my_data(data = "dbem_coords")
# List of scenarios
scenarios <- list.files(here::here("data"))
scenarios <- str_sub(scenarios,0,9)
# Run function
lapply(scenarios, grid_to_dbem)

```

## Double check grid

```{r}

read.table("~/GitHub/cocos_mpa/data/cocos_3.txt", quote="\"", comment.char="") %>%
  bind_cols(MyFunctions::my_data(data = "dbem_coords")) %>%
  filter(
    lat > -5,
    lat < 10,
    lon < -80,
    lon > -95
  ) %>% 
    # Double check matched DBEM grid (note it changes each scenario)
    mutate(
      status = ifelse(V1 == 1.5,"open",
                      ifelse(V1 < 1,"protected",
                             ifelse(V1 > 1 & V1 < 1.5,"surrouding","outside")
                      )
      )
    ) %>% 
  # View()
  ggplot() +
  geom_sf(data = ammb_sf, aes(), fill = "transparent", color = "black") +
  geom_sf(data = coco_sf, aes(), fill = "transparent", color = "red") +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = status
    )
  ) +
  scale_fill_viridis_d(alpha = 0.5)

```

## Fix scenarios

```{r}
scenario3 <- read_csv("../data/scenario3.csv")

mpa_index <- scenario3 %>% 
    rename(index = Seq) %>% 
    right_join(dbem_coords,
               by ="index") %>% 
  #   # Double check matched DBEM grid (it does)
    mutate(
      status = ifelse(prop == 1.5,"open",
                      ifelse(prop <= 1,"protected",
                             ifelse( prop > 1.5,"surrouding",
                             "outside"
                             )
                      )
      )
    ) %>% 
  # 
  # ggplot(mpa_index %>%
  #          filter(
  #   lat > -5,
  #   lat < 10,
  #   lon < -80,
  #   lon > -95
  # )
  # ) +
  # geom_tile(
  #   aes(
  #     x = lon,
  #     y = lat,
  #     fill = prop
  #   )
  # ) +
  # scale_fill_viridis_b(alpha = 0.5)
    
  ## Fix NAs for DBEM 
    arrange(index) %>% 
    mutate(prop = ifelse(is.na(prop),1.5,prop)) %>% 
    select(prop)
  
name <- paste0(here("data/cocos_3"),".txt")
write.table(mpa_index, file=name, sep="\t", col.names = F, row.names = F)


```


# Species

```{r spp_data, eval = F, echo = F}

# ETP important species
etp_species_metadata <- read_csv("~/Library/CloudStorage/GoogleDrive-jepa88@gmail.com/My Drive/lenfest_mpa_project/data/species/etp_species_metadata.csv") %>% 
  select(species,dbem) %>% 
  filter(dbem == "yes")

# DBEM species
exploited_species <- MyFunctions::my_data("sau_species") %>% 
  filter(taxon_name %in% etp_species_metadata$species) 

# Save for DBEM
write.table(exploited_species %>% 
                      select(taxon_key),
                    file = paste0(here("data","RunSppListCoco.txt")),
            sep="\t", col.names = F, row.names = F)

# Save for project
write_csv(x = exploited_species,
          here("data/project_species_list.csv"))


```
